Hydrodynamics of a Semi-Permeable Vesicle Under Flow and Confinement???

### 2020-12-15 vesicle_code/docs/permeableVes/main.tex
plot flux sign on vesicle shapes and updated notation for reduced area

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
Our analysis shows that a semi-permeable vesicle in a quiescent flow
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$. Therefore, the times for
the vesicle to deviate from its initial reduced area and to reach its
steady state are proportional to $\beta^{-1}$. This is verified for a
vesicle with initial reduced area $\nu = 0.65$ and with different
semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
to be within 1.3\% of its steady state reduced area, and the required
time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
1.3\% of its initial reduced area. Both times are proportional to
$\beta^{-1}$, and for this reason, our numerical results are reported in
terms of the scaled time $\beta t$.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

In Figure~\ref{fig:relaxationComposite}, we examine the transient
dynamics of four vesicles with $\beta = 10^{-3}$ and a variety of
initial configurations. As expected, each vesicle eventually
reaches a circular shape with reduced area $\nu=1$. Next, we observe
that even though the vesicle is inflating to a circle, meaning that
there is a net flux into the vesicle, there are regions where the flux
is out of the vesicle. These regions are denoted in blue and are only
present at early times. Finally, the vesicles's reduced area do not 
change by much at early times, and intermediate shapes of semi-permeable vesicles
resemble the shape of an impermeable vesicle. This can be explained by
considering the two terms that contribute to the vesicle velocity: (i)
the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and (ii) the
force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$, the
permeability velocity is much smaller than the force balance velocity at
early times, and the semi-permeable vesicle dynamics resembles those of
an impermeable vesicle. As the force balance velocity decreases, the
permeability velocity dominates, and the vesicle begins to inflate to a
circle with no vanishing force and velocity.

@@

We form the steady state shape for a variety of semi-permeability rates
and flow rates. Since the final shape is independent of the initial
reduce area, for all simulations we initialize the vesicle with reduced
area $\nu = 0.65$. The steady state vesicle shape and its reduced
area are in Figure~\ref{fig:shearComposite}(b). We see that the flow
rate significantly affects the final vesicle shape, but the
semi-permeability rate has a minor effect. Instead, as we saw in the
quiescent analysis, $\beta$ sets a time scale since the time required
for a vesicle to reach its steady state configuration is proportional to
$\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. (b) The steady state inclination
  angle for four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}

We consider a semi-permeable vesicle in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum imposed velocity. We examine
the effect of the non-linear Poiseuille flow on the migration of
semi-permeable vesicle. In this experiment, the vesicle is initialized
above the center line $y=0$, and it migrates towards the center
line~\cite{dan-vla-mis2009}. For an impermeable vesicle, depending on
the vesicles reduced area and the flow rate, the long time vesicle shape
is either an axisymmetric parachute or bullet, or an asymmetric
tank-treading slipper~\cite{kao-bir-mis2009}. To validate our methods,
in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
for the steady state shape of an impermeable vesicle
vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
qualitatively agree, Kaoui et al.~report that the vesicles with
$\nu=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
axisymmetric. In our simulations these vesicles are slightly asymmetric,
are undergoing a tank treading motion, and we therefore report them as
asymmetric slippers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
We consider the effect of stenosis on a semi-permeable vesicle. Inspired
by an experiment that measures the effects of cells on the dynamical
pressure-drop variations along a micrometer-sized
channel~\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to pass a vesicle
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the constriction, the result
is an increase in the pressure drop required to maintain a constant flow
rate, and the difference between these pressures is called the excess
pressure. We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\nu = 0.65$ passing through a constricted
geometry and compute the pressure drop between the inlet and outlet of
the constriction. In Figure~\ref{fig:stenosisComposite}(c), we plot the
excess pressures. The impermeable case agrees with experimental
results~\cite{abk-fai-sto2006} (see Figure 3), and the results indicate
that less pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a much longer constriction, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

@@

%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisGeom.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisGeom.pdf}
%  \caption{\label{fig:stenosisGeom} A single semi-permeable vesicle
%  passing through a constriction. The geometry and parameters are chosen
%  to replicate the experiments of Abkarian et
%  al.~\cite{abk-fai-sto2006}.}
%\end{figure}
%
%\begin{figure}[htp]
%%  \ifTikz
%%  \input{figures/stenosisPressureFluctuation.tikz}
%%  \fi
%  \includegraphics[width=\linewidth]{figures/stenosisPressureFluctuation.pdf}
%  \caption{\label{fig:stenosisPressure} The excess pressure across the
%  constriction in Figure~\ref{fig:stenosisGeom}. The black dashed lines
%  denote the start and end of the constriction. The results are
%  qualitative agreement with~\cite{abk-fai-sto2006}.}
%\end{figure}

@@

\begin{figure*}[htp]
  \centering
%  \ifTikz
  \input{figures/contractingComposite.tikz}
%  \fi
  \caption{\label{fig:contractingComposite} (a) A semi-permeable vesicle
  passing through a contracting geometry similar to the experimental
  device in~\cite{}. The narrowest part of the channel is $1.6\mu$m, and
  the vesicle height is initially $10\mu$m.}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, while its area, $A$, can change due to semi-permeability.
Therefore, the reduced area $\nu = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Quiescent Flow} 
\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/relaxationComposite.tikz}
  \fi
  \caption{\label{fig:relaxationComposite} (a): Snapshots of four
  semi-permeable vesicles with $\beta = 10^{-3}$ submerged in a
  quiescent flow. The fluid flux along the red regions is into the
  vesicle, and the fluid flux along the blue regions is out of the
  vesicle. In all cases, the vesicle reaches a steady state circular
  shape. (b): The reduced area of each of the simulations. The
  particular snapshots in (a) occur at the marks along the curve.}
\end{figure*}
Our analysis shows that a semi-permeable vesicle in a quiescent flow
relaxes to a circle at a rate that is proportional to $\beta$, and the
steady state tension is $\Lambda = 2\pi^2/L^2$. Therefore, the times for
the vesicle to deviate from its initial reduced area and to reach its
steady state are proportional to $\beta^{-1}$. This is verified for a
vesicle with initial reduced area $\alpha = 0.65$ and with different
semi-permeability rates. In Table~\ref{tbl:ellipseRelaxTimes} we report
the required time, $t_\mathrm{steady}$, for the vesicle's reduced area
to be within 1.3\% of its steady state reduced area, and the required
time, $t_\mathrm{inflate}$, for the vesicle's reduced area to exceed
1.3\% of its initial reduced area. Both times are proportional to
$\beta^{-1}$, and for this reason, our numerical results are reported in
terms of the scaled time $\beta t$.
\begin{table}[htp]
  \centering
  \caption{\label{tbl:ellipseRelaxTimes} The time for a semi-permeable
  vesicle to its steady state ($t_\mathrm{steady}$) and to begin
  inflating ($t_\mathrm{inflate}$).}
  \begin{tabular}{C{0.8cm}C{2cm}C{0.8cm}C{2cm}C{0.8cm}}
    $\beta$ & $t_\mathrm{steady}$  & ratio 
            & $t_\mathrm{inflate}$ & ratio \\
    \midrule
    $10^{0}$  & $6.45 \times 10^{0}$ & ---  
              & $2.17 \times 10^{-2}$ & --- \\ 
    $10^{-1}$ & $2.00 \times 10^{1}$ & 3.10 & 
                $1.25 \times 10^{-1}$ & 5.76 \\
    $10^{-2}$ & $9.61 \times 10^{1}$ & 4.81 & 
                $1.59 \times 10^{0}$ & 12.72 \\
    $10^{-3}$ & $8.36 \times 10^{2}$ & 8.71 & 
                $2.19 \times 10^{1}$ & 13.77 \\
    $10^{-4}$ & $8.15 \times 10^{3}$ & 9.74 & 
                $2.14 \times 10^{2}$ & 9.77 \\
    $10^{-5}$ & $8.03 \times 10^{4}$ & 9.85 & 
                $2.20 \times 10^{3}$ & 10.14 \\
    \bottomrule
  \end{tabular}
\end{table}

In Figure~\ref{fig:relaxationComposite}, we examine the transient
dynamics of four vesicles with $\beta = 10^{-3}$ and a variety of
initial configurations. As expected, each vesicle eventually
reaches a circular shape with reduced area $\alpha=1$. Next, we observe
that even though the vesicle is inflating to a circle, meaning that
there is a net flux into the vesicle, there are regions where the flux
is out of the vesicle. These regions are denoted in blue and are only
present at early times. Finally, the vesicles's reduced area do not 
change by much at early times, and intermediate shapes of semi-permeable vesicles
resemble the shape of an impermeable vesicle. This can be explained by
considering the two terms that contribute to the vesicle velocity: (i)
the permeability velocity, $\beta(\ff \cdot \nn) \nn$, and (ii) the
force balance velocity, $\SS[\ff]$. Since $\beta = 10^{-3}$, the
permeability velocity is much smaller than the force balance velocity at
early times, and the semi-permeable vesicle dynamics resembles those of
an impermeable vesicle. As the force balance velocity decreases, the
permeability velocity dominates, and the vesicle begins to inflate to a
circle with no vanishing force and velocity.

@@

We form the steady state shape for a variety of semi-permeability rates
and flow rates. Since the final shape is independent of the initial
reduce area, for all simulations we initialize the vesicle with reduced
area $\alpha = 0.65$. The steady state vesicle shape and its reduced
area are in Figure~\ref{fig:shearComposite}(b). We see that the flow
rate significantly affects the final vesicle shape, but the
semi-permeability rate has a minor effect. Instead, as we saw in the
quiescent analysis, $\beta$ sets a time scale since the time required
for a vesicle to reach its steady state configuration is proportional to
$\beta^{-1}$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Poiseuille Flow}
\begin{figure*}[htp]
  \ifTikz
  \centering
  \input{figures/parabolicComposite.tikz}
  \fi
  \caption{\label{fig:parabolicComposite} (a) The steady state shape of
  a semi-permeable vesicle submerged in a Poiseuille flow with varying
  initial reduced areas and flow rates. The fluid flux along the red
  regions is into the vesicle, and the fluid flux along the blue regions
  is out of the vesicle.  (b) The steady state inclination angle for
  four different flow velocities. The steady state shapes are
  superimposed. (c) The reduced area of the slipper formed with the flow
  rate of 800 $\mu m/s$. (d) The vesicle shape shown with the background
  imposed flow. The channel width is 12.5 times larger than the vesicle
  radius. The corresponding reduced area of each vesicle is indicated by
  the marks in plot (c).}
\end{figure*}

We consider a semi-permeable vesicle in the Poiseuille flow $\uu =
U(1-(y/W)^2,0)$, where $U$ is the maximum imposed velocity. We examine
the effect of the non-linear Poiseuille flow on the migration of
semi-permeable vesicle. In this experiment, the vesicle is initialized
above the center line $y=0$, and it migrates towards the center
line~\cite{dan-vla-mis2009}. For an impermeable vesicle, depending on
the vesicles reduced area and the flow rate, the long time vesicle shape
is either an axisymmetric parachute or bullet, or an asymmetric
tank-treading slipper~\cite{kao-bir-mis2009}. To validate our methods,
in Figure~\ref{fig:parabolicComposite}(a), we construct a phase diagram
for the steady state shape of an impermeable vesicle
vesicle~\cite{kao-bir-mis2009} (see Figure 2). While the vesicle shapes
qualitatively agree, Kaoui et al.~report that the vesicles with
$\alpha=0.7$ and flow rates $U = 1200 \mu m/s$ and $U = 1600 \mu m/s$ are
axisymmetric. In our simulations these vesicles are slightly asymmetric,
are undergoing a tank treading motion, and we therefore report them as
asymmetric slippers.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Semi-Permeable Vesicle in Stenosis}
We consider the effect of stenosis on a semi-permeable vesicle. Inspired
by an experiment that measures the effects of cells on the dynamical
pressure-drop variations along a micrometer-sized
channel~\cite{abk-fai-sto2006}, we first examining the effect of
semi-permeability on the pressure fluctuation across a stenosed
geometry. In this experiment, a pressure drop is used to pass a vesicle
red blood cell (RBC) through a constriction with diameter $5\mu$m and
length $45\mu$m. We set up a similar computational geometry with a
Poiseuille flow imposed at the inlet and outlet
(Figure~\ref{fig:stenosisComposite}(a)). The flow rate is chosen so that
the residency time of the RBC in the constriction agrees with the
experimental results. When the RBC enters the constriction, the result
is an increase in the pressure drop required to maintain a constant flow
rate, and the difference between these pressures is called the excess
pressure. We consider both an impermeable and semi-permeable vesicle
with initial reduced area $\alpha = 0.65$ passing through a constricted
geometry and compute the pressure drop between the inlet and outlet of
the constriction. In Figure~\ref{fig:stenosisComposite}(c), we plot the
excess pressures. The impermeable case agrees with experimental
results~\cite{abk-fai-sto2006} (see Figure 3), and the results indicate
that less pressure is required to maintain a constant flow rate when the
vesicle is semi-permeable. Our simulations indicate the
semi-permeability does affect the excess pressure, and the vesicle's
reduced area reduces by about 6.1\%
(Figure~\ref{fig:stenosisComposite}(d)). Since the area does decrease at
a constant rate we expect that a larger amounts of deflation is possible
if the vesicle is stenosed for a longer period of time. Therefore, in
Figure~\ref{fig:stenosisComposite}(b), we consider a vesicle passing
through a much longer constriction, and we observe that the reduced area
decreases by 13.2\% (Figure~\ref{fig:stenosisComposite}(e)).

@@

@@

Velocity scale is $100\mu$m/s.

\begin{figure*}[htp]
  \centering
  \ifTikz
  \input{figures/contractingComposite.tikz}
  \fi
  \caption{\label{fig:contractingComposite} (a) A semi-permeable vesicle
  passing through a contracting geometry similar to the experimental
  device in~\cite{}. The narrowest part of the channel is $1.6\mu$m, 
  the vesicle height is initially $10\mu$m, and the maximum background
  flow velocity is $316\mu$m/s. The fluid flux along the red regions is
  into the vesicle, and the fluid flux along the blue regions is out of
  the vesicle.}
\end{figure*}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection*{Dimensionless equations}
The inextensibility condition guarantees that the vesicle's length, $L$,
is conserved, while its area, $A$, can change due to semi-permeability.
Therefore, the reduced area $\alpha = 4\pi A/L^2 \in (0,1]$ is not
conserved. We non-dimensionalize~\eqref{eqn:vesVelocity} by introducing
the length scale $R_0 = L/2\pi$ and defining the time scale $\tau = \mu
R_0^3/k_b$, the tension scale $Q = k_b/R_0^2$, and the permeability
scale $W = R_0/\mu$. The background velocity $\uu_\infty$ is
non-dimensionalized with $R_0/\tau$.
